---
title: "Quiz-04"
author: "Shengda Hu"
author2: "Iman Chaudhry"
date: "17/10/2023"
output: 
  pdf_document: default
  html_document: default
  html_notebook: default
  slidy_presentation: default
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(carData)
```

**Before starting**

this Quiz uses the dataset `Soils` in the package `carData`. **IF** you have not installed the package `carData`, you may need to install the package `carData`. Otherwise, if you have already installed `carData` for the previous quiz, no further
installation is necessary.

### [2pts] Create dataframe

Use `tibble()` or `tribble()` to create the following dataframe (call it `NAmerCountries`), containing the top `3` most populous countries in North America:

    iso_code    location
    <chr>       <chr>
    
    USA       United States
    MEX       Mexico
    CAN       Canada

```{r}
  NAmerCountries <-
    tibble(
      iso_code = c("USA", 'MEX', "CAN"),
      location = c("United States", "Mexico", "Canada")
    )
# or
  NAmerCountries <-
    tribble(
      ~iso_code, ~location,
      #--------/----------
      "USA", "United States",
      "MEX", "Mexico",
      "CAN", "Canada"
    )
```

### Make dataframe tidy

**1)** The following two dataframes contain the same information, concerning the total COVID-19 cases and new deaths on the specific dates in a given country.

The first one is the `non_tidy` dataframe below

    date        USA           MEX         CAN         GTM
    <date>	    <chr>         <chr>	      <chr>	      <chr>
    
    2022-10-11	96783014_583	7098525_25	4306156_13	1130102_4
    2022-10-12	96847331_748	7099723_17	4309992_62	1130829_4
    2022-10-13	96911257_653	7100886_4	  4319103_109	1131666_2

The second one is the `tidied` dataframe below

    iso_code	date	        total_cases      new_deaths
    <chr>	    <date>        <int>           <int>
    
    USA       2022-10-11    96783014        583
    USA       2022-10-12    96847331        748
    USA       2022-10-13    96911257        653
    MEX       2022-10-11    7098525         25
    MEX       2022-10-12    7099723         17
    MEX       2022-10-13    7100886         4
    CAN       2022-10-11    4306156         13
    CAN       2022-10-12    4309992         62
    CAN       2022-10-13    4319103         109
    GTM       2022-10-11    1130102         4
    GTM       2022-10-12    1130829         4
    GTM       2022-10-13    1131666         2

**i)** [2pts] Use the function `read_csv` to create the two dataframes in the following code block, such that `non_tidy` contains the first dataframe in the above, while `tidied` contains the second dataframe in the above.

*Hint*: You may want to copy-paste the tables above and adapt the resulting strings to save yourself from typing everything all over.

```{r}
non_tidy <- read_csv(
  "date,CAN,GTM,MEX,USA
    2022-10-11,96783014_583,7098525_25,4306156_13,1130102_4
    2022-10-12,96847331_748,7099723_17,4309992_62,1130829_4
    2022-10-13,96911257_653,7100886_4,4319103_109,1131666_2"
)

tidied <- read_csv(
  "iso_code,date,total_cases,new_deaths
    USA,2022-10-11,96783014,583
    USA,2022-10-12,96847331,748
    USA,2022-10-13,96911257,653
    MEX,2022-10-11,7098525,25
    MEX,2022-10-12,7099723,17
    MEX,2022-10-13,7100886,4
    CAN,2022-10-11,4306156,13
    CAN,2022-10-12,4309992,62
    CAN,2022-10-13,4319103,109
    GTM,2022-10-11,1130102,4
    GTM,2022-10-12,1130829,4
    GTM,2022-10-13,1131666,2"
 )
```


**ii)** [3pts] In the code block below, write appropriate code to convert the `non_tidy` dataframe into the tidy form of `tidied`. Use the column name `cases_deaths` to store the values in the appropriate `pivot_` functions.

```{r}
tidied_by_hand <- non_tidy |>
  pivot_longer(CAN:USA, names_to = "iso_code", values_to = "cases_deaths") |>
  separate_wider_delim(
    cases_deaths,
    names = c("total_cases", "new_deaths"),
    delim = "_") |>
  relocate(iso_code)
```

**iii)** [2pts] In the code block below, write appropriate code to convert the `tidied` dataframe into the non-tidy form of the `non_tidy` dataframe. 

*Hint*: See how the dataframe is being changed by the `unite` function before working on the solution.

```{r}
non_tidy_by_hand <- tidied |>
  unite("cases_deaths", total_cases:new_deaths) |>
  pivot_wider(names_from = iso_code, values_from = cases_deaths)
```

**2)** The dataframe below is not tidy. It contain the same information as the the dataframes in the previous question, together with an extra column of `population` data.

    iso_code	total_cases	new_deaths	population
    <chr>		  <int>		    <int>		    <int>

    2022-10-11
    CAN	      4306156	    13	        38155012
    GTM	      1130102	    4	          17608483
    MEX	      7098525	    25	        126705138
    USA	      96783014	  583	        336997624
    2022-10-12
    CAN	      4309992	    62	        38155012
    GTM	      1130829	    4	          17608483
    MEX	      7099723	    17	        126705138
    USA	      96847331	  748	        336997624
    2022-10-13
    CAN	      4319103	    109	        38155012
    GTM	      1131666	    2	          17608483
    MEX	      7100886	    4         	126705138
    USA	      96911257	  653         336997624

**i)** [2pts] Use `read_csv` to create a dataframe `untidy_by_date` that contains exactly the same dataframe as above (with possible `NA`'s).

*Hint*: You may want to copy-paste the tables above and adapt the resulting strings to save yourself from typing everything all over.

```{r}
untidy_by_date <- read_csv(
  "iso_code,total_cases,new_deaths,population
    2022-10-11,,,
    CAN,4306156,13,38155012
    GTM,1130102,4,17608483
    MEX,7098525,25,126705138
    USA,96783014,583,336997624
    2022-10-12,,,
    CAN,4309992,62,38155012
    GTM,1130829,4,17608483
    MEX,7099723,17,126705138
    USA,96847331,748,336997624
    2022-10-13,,,
    CAN,4319103,109,38155012
    GTM,1131666,2,17608483
    MEX,7100886,4,126705138
    USA,96911257,653,336997624"
)
```

**ii)** [6pts] Create a new tidy dataframe `tidied_by_date` from `untidy_by_date` above in the code block below so that it has a new column `date`, recording the dates for the data contained in each row. In the result, the columns should be ordered as
`iso_code`, `date`, `total_cases`, `new_deaths`, and `population`.

```{r}
tidied_by_date <- untidy_by_date |>
  mutate(date = ifelse(is.na(total_cases), iso_code, NA)) |>
  relocate(iso_code, date) |>
  fill(date) |>
  filter(!is.na(total_cases))
```

### Read `csv` file directly

**3)** [3pts] Use appropriate command to load the `csv` file containing COVID-19 data from the following web address:

- "https://covid.ourworldindata.org/data/owid-covid-data.csv"

Then change the types of the variables whose names end with `boosters`, `cases`, `deaths`, `tests` or `vaccinations` to `integer`. 

*Hint*: Use `across` and `ends_with` within `mutate` -- do `?mutate` in the Console.

```{r}
covid <- 
  read_csv("https://covid.ourworldindata.org/data/owid-covid-data.csv") |>
  mutate(across(ends_with("boosters") | ends_with("cases") | ends_with("deaths") | ends_with("tests") | ends_with("vaccinations"), as.integer))
```
