---
title: "Quiz-03"
author: "Shengda Hu"
author2: "Iman Chaudhry"
date: "03/10/2023"
output: 
  pdf_document: default
  html_document: default
  html_notebook: default
  slidy_presentation: default
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(carData)
```

**Before starting**

You may need to install the package `carData`, since this Quiz uses the dataset `Soils` in the package `carData`.

### Basic information

[3pts] Find some basic information concerning the `Soils` data.

- Number of rows (`num_rows`)
- Number of variables (`num_cols`)
- List of categorical variables (`cat_vars`)

```{r error=TRUE}
(num_rows <-
  nrow(Soils) 
)
(num_cols <-
    ncol(Soils) 
)
(cat_vars <- 
    c("Group", "Contour", "Depth", "Gp", "Block") 
)
```

Use the Console to find out more information about the dataframe and answer the following question:

- **What does the variable `Dens` represent?**

*Answer*:
bulk density in gm/cm$^3$

### Work with the data

**I)** [8pts] Complete the code in the code blocks below to create a plot showing the covariation between the variables `Depth` and `Contour`. For each group of observations that has the same `Depth` and `Contour`, represent on the same plot the information concerning
 
+ the average pH values (create a variable `avg_pH`) and 
+ the average Calcium values (create a variable `avg_Ca`).

**Note** In the first code block, create the dataframe `computed` that contains the necessary information. In the second code block, using the resulting `computed` dataframe to complete the plot.

```{r error=TRUE}
(computed <- Soils
  |> summarize(
    .by = c(Depth, Contour),
    avg_pH = mean(pH),
    avg_Ca = mean(Ca)
  )
)
```

```{r error=TRUE}
(computed_plot <- computed
  |> ggplot(mapping = aes(x = Depth, y = Contour)) +
    geom_point(mapping = aes(color = avg_pH, size=avg_Ca))
)
```

**II)** 

**i)** [2pts] Complete the code in the code block below to create a plot showing the covariation between the two variables `Depth` and `Ca`.

```{r error=TRUE}
(cov_plot <- Soils
 |> ggplot(mapping = aes(x=Depth, y=Ca)) +
    geom_boxplot()
)
```

**ii)** [2pts] Based on the plot above, provide the name(s) of the `Depth`(s) containing no outlier in the `Ca` variable. You may want to choose appropriate method of plotting in the previous part so that it facilitates answering this question.

```{r error=TRUE}
no_outlier <- "60-90"
```

**III)** [4pts] Complete the code in the code block below to obtain the following plots of `Conduc` (on `y`-axis) against `K` (on `x`-axis):

- **1.** a scatter plot
- **2.** the line of best fit representing the linear regression model (use `lm`)

```{r error=TRUE}
(fit_plot <- Soils
 |> ggplot(mapping = aes(x=K, y=Conduc)) +
    geom_point() +
    geom_smooth(method=lm, se = FALSE)
)
```

### Descriptive statistics

**IV)** 

**i)** [4pts] Compute the covariance and correlation coefficients of the pairs of variables:

 - `K` and `Mg`
 - `Na` and `Conduc`

```{r error=TRUE}
cov_KMg <- cov(Soils$K, Soils$Mg)
cov_NaConduc <- cov(Soils$Na, Soils$Conduc) 
cor_KMg <- cor(Soils$K, Soils$Mg) 
cor_NaConduc <- cor(Soils$Na, Soils$Conduc) 
```

**ii)** [2pts] Based on the output of the code block above, describe the relationship between the pairs of variables in the dataframe `Soils`. Write your answers, with justifications, in the space provided below.

*Answer*: 
There is a relatively strong positive linear relationship between `Na` and `Conduc`, since the correlation coefficients between them is very close to `1`. There is a weak negative linear relationship between `K` and `Mg`, since the correlation coefficients between them is negative and is closer to `0` than `-1`.

**V)** [2pts] Complete the following code block to compute the correlation coefficient of the variables `Na` and `Conduc`, WITHOUT using the `R` function `cor`. Compare your answer with the output from the code block in Problem **IV)**.

```{r error=TRUE}
cor_Conduc_Na <- cov(Soils$Na, Soils$Conduc)/sqrt(var(Soils$Na)*var(Soils$Conduc)) # SOLUTION
```

**VI)** [2pts] Based on the output of the code block below, describe the relation between `Conduc` and `K` according to the linear regression model. Write down the formula expressing the relation in the space provided below.

```{r}
# nothing to change here
lm(Conduc ~ K, Soils) |>
  coef()
```

*Answer*: 
`Conduc = -13.40937 * K + 12.84066 + error`
