---
title: "Quiz-01"
author: "Shengda Hu"
author2: "Iman Chaudhry"
date: "20/09/2023"
output: 
  pdf_document: default
  html_document: default
  html_notebook: default
  slidy_presentation: default
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

This Quiz uses the dataset `txhousing`.

### Basic information

In the code block below, use one of the methods we learned to view the data in the dataframe `txhousing`.

```{r error=TRUE}
txhousing
glimpse(txhousing)
```

Then answer the questions by complete the code blocks below:

- [1pt] **Number of rows:**
```{r error=TRUE}
row_number <- 8602
```

- [1pt] **Number of variables:**

```{r error=TRUE}
variable_number <- 9
```

- [1pt] **What is the type of the variable `date`?**

```{r error=TRUE}
variable_type <- "double"
```

Use the Console to find out more information about the dataframe, and answer the following question:

- [1pt] **What do the variable `inventory` represent?**

*Answer*: 

Number of months it would take to sell all current listings at current pace of sales.

### Making plots

**I)** [2pts] Complete the code in the code block below to make a scatter plot of `sales` (on `y`-axis) against `listings` (on `x`-axis), so that the colors of the points represent `median`.

```{r error=TRUE}
ggplot(data = txhousing) +
  geom_point(mapping=aes(x=listings, y=sales, color=median), alpha=.3)
```

**II)** [2pts] Complete the code in the code block below using `date > 2010` to filter the data for a `geom_smooth` plot of `sales` (on `y`-axis) against `listings` (on `x`-axis), so that only the rows with `date` after `2010-01-01` are included.

```{r error=TRUE}
ggplot(data = filter(txhousing, date > 2010)) +
  geom_smooth(mapping=aes(x=listings, y=sales))
```

**III)** [2pts] Complete the code in the code block below so that the plots obtained in **I)** and **II)** above appear in the same plot.

**Note**: This is an exercise. In general, it's not recommended to plot different sets of data on the same graph.
```{r error=TRUE}
ggplot(data = txhousing, mapping=aes(x=listings, y=sales)) +
  geom_point(mapping=aes(color=median), alpha=.3) +
  geom_smooth(data = filter(txhousing, date > 2010))
```

**IV)** [1pts] Complete the code block below to make a box plot of `median` (on `y`-axis).

```{r error=TRUE}
ggplot(data = txhousing) +
  geom_boxplot(mapping = aes(y = median)) +
  coord_flip()
```

### Descriptive statistics

**IV)**  The values of the variable `listings` can be accessed using `txhousing$listings`.

- [2pt] Complete the code block below to find the six descriptive statistics of the variable `listings` (*Hint*: Use `txhousing$listings` to get the values of the `listings` variable as a vector). The six descriptive statistics are
    `Min`, `Q1`, `Median`, `Mean`, `Q3` and `Max`

```{r error=TRUE}
(six_stats <- 
   summary(txhousing$listings)
)
```

Write the answers in the space below. In the order of `Min`, `Q1`, `Median`, `Mean`, `Q3` and `Max`, the values are:

*Answer*:
0     
682    
1283    
3217    
2954   
43107    
1424 

- [2pt] Without using more `R` functions, find the `IQR` for the variable `listings`. Are there any outliers for the `listings` variable?

*Answer*: 

`IQR = 2954 - 682 = 2272`. 
Yes, since `Q3 + 1.5 * IQR = 2954 + 1.5 * 2272 = 5226 + 1136 = 6362 < Max = 43107`, there should be outliers above.
